### 2021-02-19 2021_realloc/02-model.Rmd
Proposition to explore new model

# A new model for commrecial data

```{r include=FALSE}
library(tidyverse)
library(ggplot2)
library(purrr)
```

As mentionned in the introduction, every  commercial catch is reported by statistical area/vessel/gear and correspond potentially to several VMS ping.
Let's denote $D_j$  a reported catch associated to $|\mathcal{P}_j|$ VMS ping at positions $\left\lbrace x_i, i\in \mathcal{P}_j \right\rbrace$. $D_j$ is the sum of all individuals catch at the $n_j$ visited fishing points:

\begin{equation}
  D_{j} = \sum_{i \in  C_j} Y_i,
  (\#eq:logbook)
\end{equation}

$C_j$ begin the vector of all fishing events associated with the logbook entry $j$.

This diffculty is currently addressed by reallocating the total catch to individual fishing events with a simple proportionality rule:
$$\tilde{Y_i} = \frac{1}{|C_j|} D_j.$$

$\tilde{Y}_i$ is the assumed to follow the observation model defined  \@ref(eq:logbook).

A more satisfying solution would consist in deriving the distribution of $D_j$ from the distribution of $Y_i$. However the sum of a mixture of delta lognormal random variables  do not resume in a nice known distribution.

An alternative would be to fit a delta lognormal model whose proportion of zero and the two first moments coincide i.e

\begin{align}
P(D_j = 0 \vert S, X) & = \prod_{i\in \mathcal{P}_j} P(Y_i = 0 \vert S, X),\nonumber \\
                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi S(x_i)}\right \rbrace}.
\end{align}

The expected catch biomass is defined by $E(D_j) = \sum_{i\in \mathcal{P}_j} E(C_i Z_i).$

As $C_i$ and $Z_i$ are assumed to be independant

\begin{align}
E(D_j) & = \sum_{i\in \mathcal{P}_j} E(C_i Z_i),\nonumber \\
& = \sum_{i\in \mathcal{P}_j} (1-p_i) \frac{\mu(x_i)}{1-p_i}, \nonumber \\
& = \sum_{i\in \mathcal{P}_j} \mu(x_i). (\#eq:newobs)
\end{align}

Regarding the variance

$$Var(D_j)  = \sum_{i\in \mathcal{P}_j} Var(C_i Z_i).$$

We are the lead to compute the Variance of $C_i Z_i$.

\begin{align*}
Var(C_i Z_i) & = E( var(C_i Z_i \vert C_i) )  + Var( E(C_i Z_i \vert C_i) ),\\
& = Var(Z_i) E(C_i^2) + Var ( E(Z_i) C_i),\\
& = Var(Z_i) (1-p_i) + E(Z_i)^2 p_i (1-p_i)
\end{align*}

## Numerical illustration

### Precising the notation 
As mentioned in Equation \@ref(eq:LNv3),  $Y\sim \mathcal{LN}(\mu, \sigma^2)$ stands  a lognormal distribution such that $E(Y) = \mu$ and $Var(log(Y)) = \sigma^2$.

An alternative  specification denotes $Y \sim LN(\rho, \sigma)$ if $E(log(Y))=\rho$ and $Var(log(Y)) = \sigma^2$

The code below is an helper to go back and forth between the different parametrizations.
```{r lognomral_moments}

LN_variance_v1 <- function(rho, sigma){
  ( exp(sigma^2) - 1 ) * exp(2* rho  + sigma^2)
}

LN_mean_v1 <- function(rho, sigma){
   exp( rho + sigma^2/2 )
}

LN_rho2mu <- function(rho, sigma){
   LN_mean_v1(rho, sigma)
}

LN_mu2rho <- function(mu, sigma){
   log(mu) - sigma^2 /2
}

LN_variance_v3 <- function(mu, sigma){
  mu^2* (exp(sigma^2)-1)
}

```

### The sum of lognormal random variables is reasonnably log normal

#### n = 2

```{r sum_lognormal}
set.seed(1234)
n <- 2
rho <- rnorm(n, mean = 0, sd = 1 )
sigma <- 01
```

We are ready to look at the sum of `r n` lognormal random variables with mean `r round(LN_mean_v1(rho, sigma),3)`  and variance `r  round(LN_variance_v1(rho, sigma),3)` .

```{r sample_2_lognorm, echo = FALSE}
n_sim <- 1000
dta <- rho %>% map(function(x){ exp(rnorm(n = n_sim, mean = x, sd = sigma))}) %>% 
   map_df(
         ~ tibble(x = .x), 
         .id = "dist"
     )

sum_dta <- dta %>% group_by(dist) %>% 
  mutate(id = row_number()) %>% 
  pivot_wider(names_from = dist,    names_glue = "dist_{dist}",
 values_from =x) %>% 
  mutate( D = rowSums(.[grep("dist_", names(.))]))

p <- sum_dta %>% ggplot() + geom_histogram(aes(x=D, y = ..density..))

mu_sum <- sum(LN_rho2mu(rho, sigma))
S2_sum <- sum(LN_variance_v1(rho, sigma))
sigma_sum = sqrt( log(S2_sum/mu_sum^2 +1))

sum_density_LN <- tibble(x  = seq(0.001, max(sum_dta$D)), 0.1) %>% 
  mutate( log_x= log(x), dens = dnorm(log_x, 
                         mean=LN_mu2rho(mu_sum, sigma_sum), 
                         sd = sigma_sum)/x)

p_LN <- p + geom_line(data= sum_density_LN, aes(x=x, y =dens), col = 'blue')
sum_density_N <- tibble(x  = seq(0.001, max(sum_dta$D), 0.1)) %>% 
  mutate(dens = dnorm(x, 
                         mean=mu_sum, 
                         sd = sqrt(LN_variance_v3(mu_sum, sigma))))

p_N <- p_LN + geom_line(data= sum_density_N, aes(x=x, y =dens), col = 'red')
p_N
```

### n = 5

```{r sum_lognormal_n5}
n <- 5
rho <- rnorm(n, mean = 0, sd = 1 )
sigma <- 01
```

We are ready to look at the sum of `r n` lognormal random variables with mean `r round(LN_mean_v1(rho, sigma),3)`  and variance `r  round(LN_variance_v1(rho, sigma),3)` .

```{r sample_2_lognorm_n5, echo = FALSE}
n_sim <- 1000
dta <- rho %>% map(function(x){ exp(rnorm(n = n_sim, mean = x, sd = sigma))}) %>% 
   map_df(
         ~ tibble(x = .x), 
         .id = "dist"
     )

sum_dta <- dta %>% group_by(dist) %>% 
  mutate(id = row_number()) %>% 
  pivot_wider(names_from = dist,    names_glue = "dist_{dist}",
 values_from =x) %>% 
  mutate( D = rowSums(.[grep("dist_", names(.))]))

p <- sum_dta %>% ggplot() + geom_histogram(aes(x=D, y = ..density..))

mu_sum <- sum(LN_rho2mu(rho, sigma))
S2_sum <- sum(LN_variance_v1(rho, sigma))
sigma_sum = sqrt( log(S2_sum/mu_sum^2 +1))

sum_density_LN <- tibble(x  = seq(0.001, max(sum_dta$D)), 0.1) %>% 
  mutate( log_x= log(x), dens = dnorm(log_x, 
                         mean=LN_mu2rho(mu_sum, sigma_sum), 
                         sd = sigma_sum)/x)

p_LN <- p + geom_line(data= sum_density_LN, aes(x=exp(log_x), y =dens), col = 'blue')

sum_density_N <- tibble(x  = seq(0.001, max(sum_dta$D), 0.1)) %>% 
  mutate(dens = dnorm(x, 
                         mean=mu_sum, 
                         sd = sqrt(LN_variance_v3(mu_sum, sigma))))

p_N <- p_LN + geom_line(data= sum_density_N, aes(x=x, y =dens), col = 'red')
p_N
```

The expected value equals  `r mu_sum`, while the mean equals  `r mean(sum_dta$D)`. The variance equals `r LN_variance_v3(mu_sum, sigma_sum)` and the empiric variance equals `r var(sum_dta$D)`.

### What happens with the addittion of some zero inflation