### 2022-03-30 Version_M2_plusieurszones/draft/paper_3/3_formulas.Rmd
paper_3

---
title: "**Formulas for the model built on catch declarations**"
output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Notations**

We model catch declarations $D_j$ (available at coarse resolution through logbooks data) as a sum of $Y_{i}$ punctual observations (which are unknown i.e. latent variables) each one realised at one fishing position $x_{i}$ (known through VMS data).

We note:

- $\mathcal{P}_j$ : the vector of all fishing positions related to the $j^{th}$ declaration.

- $j \in \{1,...,l\}$ with $l$ the number of declarations.

- $i \in \{1,...,m_j\}$ with $m_j$ the number of fishing positions belonging to the $j^{th}$ declaration.

$$D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}$$

**Reparameterization of the lognormal distribution**

The lognormal distribution is usually written as $Z \sim \mathcal{LN}(\rho;\sigma^2)$ with $Z=e^{\rho+\sigma N}$ and $N \sim \mathcal{N}(0,1)$. In this case, $E(Z)=e^{\rho + \frac{\sigma^2}{2}}$ and $Var(Z)=(e^{\sigma^2}-1)e^{2 \rho + \sigma^2}$.

We choose to slightly reparameterize the lognormal distribution. Let's define $\rho = ln(\mu) - \frac{sigma^2}{2}$, then:

- $Z=\mu e^{\sigma N - \frac{sigma^2}{2}}$

- $E(Z)=\mu$

- $Var(Z)=\mu^2(e^{\sigma^2} - 1) \Leftrightarrow	\sigma^2=ln(\frac{Var(Z)}{E(Z)^2} + 1)$

**$D_j$ probability distribution and moments**

We have to express the probability distribution of $D_j$ and its moments as a function of $Y_{i}$ and its related moments. Let's assume $Y_{i} = C_{i}.Z_{i}$ is a zero-inflated lognormal distribution with $C_{i}$ a binary random variable and $Z_{i}$ a lognormal random variable.

$$C_{i} \sim \mathcal{B}(1-p_{i})$$ 
with $p_{i}=exp(-e^\xi .S(x_{i}))$ the probability to obtain a zero value.

$$Z_{i} \sim \mathcal{LN}(\frac{S(x_{i})}{1-p_{i}},\sigma^2)$$

Here, $Y_{i}$, $C_{i}$ and $Z_{i}$ are observations of a latent field $S(x_{i})$ at a sampled point $x_{i}$.

**Probability of obtaining a zero declaration**

\begin{align*}
P(D_j = 0 \vert S, X) & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0 \vert S, X),\nonumber \\
                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j.
\end{align*}

**Expectancy of a positive declaration**

Following calculations are supposed to be conditionnal on $S$ and $X$.

$$E(D_j\vert D_j >0) =  \sum_{i\in \mathcal{P}_j} E(C_{i} Z_{i}\vert \exists i\in\mathcal{P}_j , C_{i}=1)$$

\begin{align*}
E(D_j\vert D_j > 0) & = E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } )  / P\left ( D_j > 0\right ),   \\
& = E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } )  / \left (1-\pi_j\right).   \\
\end{align*}

As $E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } ) = E(D_j ),$

\begin{align*}
E(D_j\vert D_j > 0) & = \left (1- \pi_j\right)^{-1} E(D_j)   ,   \nonumber \\
& = \left (1- \pi_j\right)^{-1} \sum_{i\in \mathcal{P}_j} E(C_{i} Z_{i}),\nonumber \\
& = \left (1- \pi_j\right)^{-1} \sum_{i\in \mathcal{P}_j} (1-p_{i}) \frac{S(x_{i})}{1-p_{i}}, \nonumber \\
& = \left (1- \pi_j\right)^{-1}\sum_{i\in \mathcal{P}_j} S(x_{i}).
\end{align*}

**Variance of a positive declaration**

$$Var(D_j \vert D_j >0)  = E(D_j^2 \vert D_j >0)- E(D_j \vert D_j >0)^2.$$

$$E(D_j^2 \vert D_j >0) = (1-\pi_j)^{-1} E(D_j^2 1_{\left \lbrace D_j >0\right\rbrace}) = (1-\pi_j)^{-1} E(D_j^2 )$$

$$E(D_j \vert D_j >0)^2 = ((1-\pi_j)^{-1} E(D_j 1_{\left \lbrace D_j >0\right\rbrace}))^2 =  (1-\pi_j)^{-2} E(D_j)^2$$

Then,

$$Var(D_j \vert D_j >0) = (1-\pi_j)^{-1} E(D_j^2 ) - (1-\pi_j)^{-2} E(D_j)^2 = (1-\pi_j)^{-1} Var(D_j) - \frac{\pi_j}{(1-\pi_j)^2} E(D_j)^2.$$

As the $(Y_{i})_{i\in \mathcal{P}_j}$ are independent, $Var(D_j) =\sum_{i \in\mathcal{P}_j} Var(Y_{i}) = \sum_{i \in\mathcal{P}_j} Var(C_{i}.Z_{i})$.

\begin{align*}
Var(C_{i} Z_{i}) & = E( C_{i}^2 Z_{i}^2 ) - E(C_{i} Z_{i})^2,\\
& =  E( C_{i}^2) E(Z_{i}^2 ) -E(C_{i})^2 E(Z_{i})^2,\\
& = (1-p_{i}) E(Z_{i}^2 ) - (1-p_{i})^2 E(Z_{i})^2,\\
& = (1-p_{i})(Var(Z_{i})+E(Z_{i})^2)-(1-p_{i})^2E(Z_{i})^2,\\
& = \frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-1)+\frac{S(x_{i})^2}{1-p_{i}}-S(x_{i})^2,\\
& = \frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i}))
\end{align*}

**Sum up of the main formulas**

$$P(D_j = 0 \vert S, X) = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j$$

$$E(D_j \vert D_j > 0)=\frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}$$

$$Var(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} Var(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2$$

$$Var(Y_{i})=\frac{S(x_{i})^2}{1-p_{j}}(e^{\sigma^2}-(1-p_{i}))$$

Assuming $D_j \vert D_j > 0$ also follows a lognormal distribution we can write:

$$D_j \vert D_j > 0 \sim \mathcal{LN}( \mu_j = E(D_j \vert D_j > 0), \sigma_j^2= ln(\frac{Var(D_j \vert D_j > 0)}{E(D_j \vert D_j > 0)^2} + 1))$$

### Simulations

```{r,echo=F,fig.asp=1,fig.width=8,fig.height=2,fig.align='center'}

obs_model <- "zerolognorm"
q1 <- 2 # zero-inflation parameter (when increases, nb of zero increases)
SD_obs <- 0.1  # observation error (standard deviation for lognormal distribution, CV for gamma)
n_obs <- 10 # number of observations per fishing sequence (number of observations over which the catches are summed)
n_sim <- 1000 # number of replicates for each 'mu' value
source("r/3_simulations.R")

```