### 2022-05-04 Version_M2_plusieurszones/draft/paper_3/paper_3.Rmd
WIP

---
title             : "Downscaling coarse observations to predict continuous wild species spatial distribution"

author: 
  - name          : "Author1"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Author2"
    affiliation   : "1,2"
    role:
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "Instit1"
  - id            : "2"
    institution   : "Instit2"

authornote: |
  Test

abstract: |
  Test

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["paper_3.bib"]

floatsintext     : yes
figurelist       : no
tablelist        : no
footnotelist     : no
linenumbers      : yes
mask             : no
draft            : no

documentclass     : "apa6"
classoption       : "man"
header-includes:
   - \usepackage{graphicx}
   - \usepackage{caption}

output            :
  papaja::apa6_pdf:
    includes:
            in_header: header.sty
    # reference_docx: mystyles_number_test.docx
---

```{r setup, include = FALSE}
library(cowplot)
library(doBy)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(kableExtra)
library(knitr)
library(latex2exp)
library(mapdata)
library(scales)
library(papaja)
library(rnaturalearth)
library(sf)
```

```{r analysis-preferences}

# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed, cache = TRUE)

mapBase <- map("worldHires", fill = T, plot = F)
mapBase <- st_as_sf(mapBase) %>% filter(ID %in% c("France","Spain"))
```

# Introduction

**Attention:** L'introduction n'est pas finie

**Context**

Spatial ecological data are becoming more and more accessible every days through new technologies and thanks to the huge effort from the scientific community to provide intensive information for ecology, evolution and conservation [@nathan2022big;@hampton2013big;@gremillet2022big]. These data are crucial to face the current challenges related to large and small scales ecological questions: for instance, tracking climate change [@maureaud_are_2020], following animal movement [@nathan2022big] or mapping species distribution [@isaac2020data]. In particular, a key point of these studies lies in understanding the drivers of the ecological processes under study. For instance, when investigating species distribution one will try to find the main covariates (e.g. substrat, temperature, depth/altitude) that shape species distribution [@guisan_predictive_2000;@planqueUnderstandingWhatControls2011].

These datasources are often highly heterogeneous in size, type and sampling design making there combination a methodological challenge [@pacifici2017integrating;@fletcher_practical_2019;@miller2019recent;@renner2019combining;@isaac2020data]. For instance, in species distribution modelling, recent studies have investigated how to combine scientific standardized data with citizen science data as they prooved to be highly complementary. The first one benefit from standardized protocol, controlled sampling plan and is usually considered as high-quality data. The second one provide a larger amount of data with lower cost, but is also caracterized by misreporting or opportunistic sampling and then by potential sampling bias [@botella2021jointly;@steen2021spatial].

Another less frequently mentionned data are declarations data (we refer to declaration data as the mandatory data that must be reported by some agent as a legal requirement to proceed with his activity). As they are mandatory, declarations data are usually very large datasets and thus they may be very informative of species spatial ecology. A common example of such datasources are commercial catch declarations data in fisheries science [@hintzenVMStoolsOpensourceSoftware2012] or harvest data in hunting [@bauder2021mismatched;@gilbert2021integrating]. These data are most often declared at the scale of coarse spatial units while scientific survey (and commonly citizen science data) are usually reported with their exact locations. Furthermore, these administrative units do not generaly have a resolution that is relevant for ecological analysis and conservation [@pacifici2019resolving].

Consequently (in ecological applications but in other fields too), to bypass the difference of resolution it is standard to apply rough hypothesis through geoprocessing technics based on proportional allocation and centroid smoothing [@gotway2007geostatistical;@young2007linking;@hintzenVMStoolsOpensourceSoftware2012]. However, these methods do not provide a measure of uncertainty related to the processing of the data. Overall, consequences of such procedures are hard to predict and can result in strong and deleterious biases [@gotway2007geostatistical;@pacifici2019resolving]. These bias do not only affect the mean and the variance, but also any statistic that would be derived from these estimates. Developing statistical methods to properly infer spatial ecological process at a fine scale from coarse declarations data and integrate these with higher resolution data is then a challenge that could give access to huge amount of information.

**The COS issue**

Reconciling spatial (and eventually temporal) scales properly when different set of observations do not have the same resolution is a well known issue in geography, ecology, agriculture, geology and statistics [@gotway2002combining]. Inferring values of a variable at points or blocks different from those that have been observed is commonly refered as the *change of support problem* (COS) in the statistical literature [@gotway2002combining and @gelfand2010misaligned]. COS issues can be classified within 2 categories - that are not necessarily mutually exclusive - [@pacifici2019resolving]:

- (1) 'spatial and temporal misalignement' (the case of some multivariate data where not all variables have been observed over all locations - see for example @finley2014bayesian).

- (2) 'modifiable areal unit' problem (MAUP) when data are aggregated over increasingly larger geographic scales (e.g. collected at point level but regrouped/declared at coarse level - see for instance @gilbert2021integrating). Our case of application clearly refers to MAUP. Note that within this category, another category exist and is commonly refered to the 'ecological fallacy'. This occures when the individual response to a covariate differs from the response estimated from grouping individuals. This means that conclusions from a fine-resolution analysis can differ from an analysis at a coarser scale based on the aggregation of the fine-resolution data.

Since 2000, several studies have described how could be overcome COS issues; @mugglin2000fully, @gelfand2001change, @gotway2007geostatistical, @wikle2005combining proposed generic approaches (and extensions of these approaches - @kim2016change) for adressing COS in spatial or spatio-temporal context. In health analysis, @young2007linking proposed to compare some standard (and rough) way to connect data that do not have the same spatial resolution with a method properly accounting for COS. @berrocal2010bivariate and @berrocal2010spatio proposed a spatio-temporal method for fusing several air pollution data: one from coarse resolution but with full spatial coverage and another recorded at point level, with sparse distribution but where records almost corresponds to the true value of the process. In climate science, @reich2014spectral and @parker2015multiresolution proposed a spectral statistical approach to downscale information from large scale model to lower scales. In the field of ecology, some recent studies have tackled such issues: @finley2014bayesian provided a framework for integrating spatially misaligned data, @hefley2017bias proposed a solution based on COS to account for location error in presence-only data, @pacifici2019resolving introduced a framework for integrating datasources of different resolution to map species distribution. Applying similar ideas, @gilbert2021integrating integrated harvest data (aggregated data) and camera trap (punctual data) to map several wildlife species in Wisconsin.

**Focus of the paper**

Still, applications in ecology remains sparse in regards to the huge amount of data that is now available. One of the limiting point consist in the kind of observational data that can be fitted to the existing COS framework. Indeed, the framework that were developped so far and their related applications mainly limited their scope to relatively simple observational data: count data were modelled through Poisson processes [@mugglin2000fully;@gotway2007geostatistical;@pacifici2019resolving;@gilbert2021integrating] and continous data were modelled through Gaussian or Gamma distributions [@gelfand2001change;@wikle2005combining;@berrocal2010spatio]. However, ecological data do not always consist of observations that can be modelled with standard probability distributions. For instance, in frequent cases data may be zero-inflated and positive-continous data. Several studies have developed models to handle properly such data in a computationaly efficient way [@lecomte2013compound;@thorson2018three]. However, these may complicate a bit the way COS is tackled when dealing with an aggregation of such complex data as their convolution may not be as simple as Poisson or Gaussian ones.

In this paper, we aim at illustrating how to deal with change of support in ecological applications when the observational data is complex (e.g. zero-inflated and positive continous). We develop our approach on an existing framework developped by @alglave_combining_2022 in the field of marine and fisheries ecology. The framework aims at predicting the spatial distribution of fish species based on 2 datasets: scientific survey data and commercial catch declarations data. Commercial catch declarations are declared at the level of ICES rectangles (resolution of 0.5° x 1°) while scientific data benefit from exact location records. Usually in standard processing, declarations data are reallocated uniformly over their GPS fishing positions (available through Vessel Monitoring Satellites - VMS) in order to improve its spatial resolution.

In the following, we first describe the original model integrating both datasources (part 1). We then propose a generic statistical solution that allow to properly tackle the change of support issue and adapt it to our specific case (part 2).

Finally, we compare this method to the current geoprocessing method in 3 steps:

- A first simulation-estimation step at the scale of a single statistical rectangle. Our aim is to explore the base properties of our framework and then the biomass distribution is simplified to the most basic version (part 3).

- A second simulation-estimation step at the scale of several statistical rectangles. The latent field is complexified and the set of simulations is parameterized so as to get closer to a realistic case study (part 4).

- A final step where we compare the 2 methods on a real case study (common sole in the Bay of Biscay) and illustrate the main results that are consitent with simulations-estimations (part 5).

#  Introduction 
* Commencer par la motivation peche
--> knowledge of essential habitat --> Species distribution model --> link between biomass and habitat to identify essential habitat
-->  Data from scientific survey, precisely geolocalized standard sampling protocol, expensive and time consuming
--> On the other hand declaration data, spatialized at coarser level  but cheap and reglementaire
Combining the two sources of data.

This problem is not unique, other exampl in terms of apllication

Change of support form fine to coarse sptial level

A solution the reallocation process

** Pour la discu, simulation conditionnelle ? **

We proposed an alternative approach to account for the change of support and we compare it to the approach proposed in alglave but also to a golden standard to assess the consequences of these methods.

Paper organized as follow

Section 2. presentation of the reference model and detail the reallocation process

Section 3. Presnetation of an alternaion way to deal with change of support

Section 4. simulation study to asses the bias and/or the loss of precision of the two methodes compared with the golden standard

Section 5. Appllication on sole of the Bay of Biscay

Section 6. Discu/conclu

## An explicitly spatialized catch model

In  @alglave_combining_2022, the authors propose a hierarchical spatial model to combine scientific survey data obtained through a standardized sampling protocol and catch data harvested by fishermen. This section provides a biref overview of the key ingredient of this model and raises some concern on the spatialization of the commercial catch data.

Let  $D\subset \R^2$ be a spatial domain and $\bm{S}=(S(x), x\in D)$ a spatial random field which represents the biomass for a species of interest.  $\bm{S}$ is assumed  to be a spatial log-Gaussian Random Field (GRF) defined as $\log(S(x))=\mu + \beta \, \Gamma(x) + \delta(x)$ (Figure \@ref(fig:SchemeRealloc)) where $\bm{\delta}=(\delta(x), x\in D)$ is a zero mean GRF with covariance matrix $\Sigma$.  As the catch are   zero-inflated positive continuous data,  following  @thorson_three_2018,  the authors model a catch at a given site $X_i$ with a mixture of a Dirac mass at $0$ and a Log Normal distribution, the proportion of the mixture being defined  conditionally on the random field $\bm{S}$.

\begin{align}
Y_i | S(x_i) & \overset{ind}{\sim} p_i \delta_0+ (1-p_i) B_i, \quad B_i \sim \mathcal{L}\left ( \frac{S(x_i)}{1-p_i} ,\sigma^2\right), \cr \nonumber
\mbox{with } p_i &= \exp(-e^\xi S(x_{i}))
\end{align}

<!-- 
Je pense qu'il ne faut pas parler de la grille discrete tout de suite car S est défini sur un espace continu et moralement l'intensité du processus de poisson aussi. le passage à la grille est la pour simplifier l'expression de la vraisemblance, on peut dire qu'on fait une approximation pour des raisons numériques

We consider a discrete study domain where $x \in [\![1,n]\!]$ stands for the discrete locations. $\beta . \Gamma(x)$ is the covariate term with $\Gamma$ the matrix of design of the covariates and $\beta$ the effect of the covariates. $\delta(x)$ is a GF capturing spatial correlation ($\delta \sim \mathcal{N}(0,\Sigma)$). -->

In this approach all fishing locations $x_i$ and the coressponding catch are supposed to be known.

<!-- however those fishinf locations are actually derived from a different dataset containing the hourly localisation of the the commercial vessels, the VMS dataset. This dataset is processed to identify the fishing location among all the recorded locations. -->

<!-- At each fishing location $x_i$, fishermen realize a catch $Y_i$ conditionnally on the latent field values (referred hereafter as a punctual catch). Observations are supposed to follow some distribution $\mathcal{L}_Y$. -->

<!-- $$Y_i | S(x_i), x_i \sim \mathcal{L}_Y(S(x_i),\xi,\sigma^2)$$ -->

<!-- In our case, observations are zero-inflated positive continuous data and they are modelled through a zero-inflated lognormal model previously introduced by @thorson_three_2018 and already used in @alglave_combining_2022. $\mathcal{L}_Y$ is decomposed it 2 parts: -->

<!-- - the probability to obtain a zero-catch which is modelled as a Bernoulli variable with probability $p_{i}=\exp(-e^\xi .S(x_{i}))$. $\xi$ is a parameter controlling zero-inflation. When $\xi$ increases, the amount of zero in the data decreases. -->

<!-- - if the catch is positive, the probability to obtain some observation value $y_i$ is modelled through a positive continous distribution $\operatorname{L}$ (here a lognormal distribution - its parameterization is described in SM)) with mean component $\frac{S(x_{i})}{1-p_{i}}$ and variance term $\sigma^2$. -->

<!-- $$ -->
<!-- \operatorname{P}\left(Y_i=y_{i} | x_i, S(x_i)\right) = -->
<!-- \left\{ -->
<!--     \begin{array}{ll} -->
<!--         p_i & \text { if } y_{i}=0 \\ -->
<!--         \left(1-p_i\right) \cdot \operatorname{L}\left(y_{i },\frac{S(x_i)}{\left(1 - p_i\right)},\sigma^{2} \right) & \text { if } y_{i} > 0 -->
<!--     \end{array} -->
<!-- \right. -->
<!-- $$ -->

Actually, fishermen do not have to declare this precise information, but it is mandatory, for every vessel to declare the total daily catch aggregated at a given administrative spatial unit named  statistical rectangle. These data are recorded in logbooks data. For a given vessel, on a given day, a declaration (denoted $D_j$) is therefore the sum of all individual catch $Y_i$ realized in the unit associated with the catch  declaration $D_j$, more formally:

$$D_j=\sum_{x_i \in \mathcal{A}_j}{Y(x_i)},$$

where $\mathcal{A}_j$ is the administrative spatial unit associated to declaration $j$.

The original model proposed in  @alglave_combining_2022 is defined at the individual catch scale, those catch are  derived from the catch declaration combined with the VMS data. This is classically done in fisheries science REFREFEREFREF by using a uniform reallocation process. This process consists in counting, for a given vessel, on a given day,  the number $m_j$ of fishing points in $\mathcal{A}_j$ and define for each $x_i\in\mathcal{A}_j,$, the associated individual catch $Y(x_i)=D_j/m_j$. As a direct consequence, the so reconstructed individual catch are similar to a Catch Per Unit of Effort (CPUE), effort being measure in hours. As noted by @alglave_combining_2022, this process has several drawbacks.
First, as a consequence of the reallocation process, the so reconstructed individual catches tend to exhibit smoother pattern than the original catches. Second, the actual number of data is the total number of declaration while the number of data after the reallocation process is the number of fishing locations, which is approximately 10 times the number of declarations. From a statistical point of view, this overestismation of the number of informative data tends to produce excessively narrow confidence intervals.

<!-- **Catch reallocation: uniform reallocation or model-based reallocation** -->

<!-- In standard processing, $D_j$ (available from logbooks) are reallocated uniformly on related $x_i$ (available from VMS) so that derived punctual observations $Y^*_i$ are computed as $Y^*_i=D_j/m_j$. This is what we call uniform (or proportional) reallocation. In this case, inference of species distribution is directly computed through reallocated $Y^*_i$ assuming these are the exact punctual observations. This strongly simplifies the actual process of observation and have most likely repercussion on model performance. -->

<!-- To overcome such limitations, an alternative is to consider that only the catch declarations $D_j$ are observed while $Y_i$ are not (they are latent variables). Such way, we define some distribution $\mathcal{L}_D$ as the distribution of the catch declarations: -->

To circumvent such limitations, we propose an alternative approach that models the catch declarations $D_j$ instead of the individual catch $Y(x_i)$. As we  aim to propose a model compatible with the original one, we are lead to specify  the distribution of $D_j$ which consists of a sum of $m_j$ random variables folowwing a mixture distribution. This distribution has no  known analytical form. However the catch data also exhibit some zero inflation and a skewed repartition of the values and a mixture model would be a good candidate.  Theefore, we propose to model $D_j$ as a mixture between a Dirac mass in $0$ and a lognormal distribution. In order to compare the original approach with our proposition, we define the different key quantities, i.e  the mixture proportion, the expected positive catch declaration and its variance so that to match the original propostion:

<!-- that only the catch declarations $D_j$ are observed while $Y_i$ are not (they are latent variables). Such way, we define some distribution $\mathcal{L}_D$ as the distribution of the catch declarations: -->

$$D_j | \mathcal{P}_j,\bm{S} \sim \mathcal{L}_D( \mu_j,\xi,\sigma^2)$$

with $\mathcal{P}_j=(x_1,...,x_i, ...,x_{m_j})$, the list of the fishing points associated to declaration $D_j$,  $\mu_j$ the expected positive biomass, $\xi$ the zero-inflation parameter and $\sigma^2$ the variance parameter.

<!-- As $D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}$, it is possible to relate $D_j$, $Y_i$ and $S_{\mathcal{P}_j}$ through the distribution $\mathcal{L}_Y$ and $\mathcal{L}_D$. Our approach is to match the moments of $D_j$ (obtained from the moments of $\mathcal{L}_Y$) and $\mathcal{L}_D$ and then to infer $S(x)$ and $Y_i$ from $D_j$. This is what we call model-based reallocation. This model will be referred as the $D_j$ model. In contrast, the model fitted on reallocated observations $Y^*_i$ will be referred as the $Y^*_i$ model. -->

<!-- In the following, $Y$ and $D$ will be assumed conditionnal on the fishing positions and the latent field values, but for notation simplicity they will be simply denoted as $Y_i$ or $D_j$ instead of $Y_i \vert x_i, S(x_i)$ and $D_j | \mathcal{P}_j,S_{\mathcal{P}_j}$. -->

**Matching $\mathcal{L}_Y$ and $\mathcal{L}_D$**

First, as at the punctual level ($Y$), the declarations can be decomposed in 2 components: (1) the probability to obtain a zero-value declaration and, if the declaration is positive, (2) the probability to obtain some declaration value $d_j$.

(1) We derive the probability to obtain a zero-declaration $P(D_j = 0)$ by simply multiplying the probability to obtain a zero-punctual observation $P(Y_i=0 )$ to all fishing points $x_i \in \mathcal{P}_j$.

\begin{align*}
P(D_j = 0) & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0),\nonumber \\
                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j.
\end{align*}

(2) If the declaration is positive, we assume the probability to obtain catch $d_j$ follows a continous distribution $\mathcal{L}_{D_j|D_j>0}$ with mean component $\mu_j$ and variance component $\sigma_j$. From $\mathcal{L}_Y$ we can derive $E(D_j \vert D_j > 0)$ and $Var(D_j \vert D_j > 0)$ (given below) and re-express $\mu_j$ and $\sigma_j$ as a function of these two moments.

$$E(D_j \vert D_j > 0)=\frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}$$
$$Var(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} Var(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2$$
$$Var(Y_{i})=\frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i}))$$

If we consider $\mathcal{L}_{D_j|D_j>0}$ is lognormal (with parameterization described in SM) then, $\mu_j = E(D|D_j>0)$ and $\sigma_j = ln(\frac{Var(D|D_j>0)}{E(D|D_j>0)^2} + 1)$. Such way, by fitting the model on declaration $D_j$ rather than reallocated catch $Y_i$ and linking $\mathcal{L}_D$ to $\mathcal{L}_Y$, we intend to better follow the observation process of the commercial data and potentially improve inference compared with the approximation realized through uniform reallocation. All calculations to obtain these formulas are available in SM.

The inference is realized through Template Model Builder (TMB), an effective tool to build hierarchical models and perform maximum likelihood estimation through automatic differentiation and Laplace Approximation [@kristensen_tmb_2016].

![(\#fig:SchemeRealloc) Schematic representation of the reallocation process. The biomass field (the background field) depends on a covariate and a spatial random effect. The covariate is the $x$ axis. It has a positive effect on biomass values (i.e. biomass is higher on the right of the grid than on the left). The spatial random effect conduct to a hotspot on the bottom-right of the latent field. The study domain is considered as a statistical rectangle (grey square). Fishermen sample catches in areas of poor biomass where the covariate is relatively low (blue points) and in areas of higher biomass where the covariate is higher and eventually in the hotspot of biomass (orange and red points). These catches belong to the same declaration $j$ and are summed to constitute the declaration $D_j=50$. The declaration is declared at the level of the statistical rectangle. From VMS data, we know the fishing positions $x_i$. In standard processing, $D_j$ are then uniformly reallocated over the fishing positions $x_i$. This strongly homogeneizes the catch. In particular, the effect of the habitat is no more evidenced in the reallocated catch $Y_i^*$.](images/realloc.png)

```{r}

# save(data=gridpolygon_sf_2,file="res/simu_lf.RData")
# 
# plot_1 <- ggplot(gridpolygon_sf_2)+
#   geom_sf(aes(fill=Strue_x),alpha=0.9)+
#   scale_fill_distiller(palette = "Spectral")+
#   theme_void()+theme(legend.position = "none")
# 
# ggsave("images/simu_lf.png",width = 7.5,height = 7.5,dpi = 500)

```

## Simulation-estimation

To evaluate the impact of reallocation on model outputs and how our alternative approach can improve model accuracy, we conduct simulation-estimation.

First, we conduct basic simulations in order to explore the effect of some key variables on model performance and also the base properties of the alternative models. We assume the study domain is a single statistical rectangle, only commercial data feed the model and the expression of the latent field in both simulation and estimation is simplified (the variability only depends on one covariate, not on a spatial random term). These simulations will be referred as "single-square simulations".

Then, we extend the simulation-estimation study to several rectangles to get closer to a real case study configuration. We add a spatial random effect in the latent field (for both simulation and estimation) and we also simulate scientific data (in addition to commercial data) to explore the contribution of both datasets in inference. These simulations will be referred as "multiple-square simulations".

In these two sets of simulation-estimation, the covariate is modelled as a continous GRF that we suppose known at each point of the grid. The covariate effect is fixed to $\beta_S=2$ and the intercept is also fixed to $\mu=2$. Regarding commercial data, the number of fishing pings per declaration is fixed to 10 as it is the average number of fishing locations for a single declaration in real data. All parameterizations are detailed in the Table 1.

Each time we compare several model configurations and evaluate these in regards to 2 metrics:

- the MSPE which quantifies the accuracy of the spatial predictions of the latent field over the spatial domain ($n$ is the number of locations over the grid).

$$MSPE=\frac{\sum_x (S(x) - \hat{S}(x) ) ^2}{n}$$

- the parameter of the species habitat relationship $\beta_S$ compared with its estimate $\hat \beta_S$.

```{r}

df <- data.frame(v1=c("$\\mu$","$\\beta_S$","$\\text{Range}$","$\\text{Marginal variance}$","$\\xi_{com}$","$\\sigma_{com}$","$k_{com}$","$\\xi_{sci}$","$\\sigma_{sci}$"),
                 v2=c("2","2"," -- "," -- ","-1","1"," -- "," -- "," -- "),
                 v3=c("2","2","1.5","0.5","-1","1","1","0","0.8"))

# italic column headers
colnames(df) <- c("\\text{Parameters}","\\text{Single-square simulations}","\\text{Multiple-square simulations}")

kable(df, "latex", align=c("l","c","c"), booktabs=TRUE, escape = F, caption = 'Parameter values for the simulations')

table_param <- t(data.frame(
  c("$\\mu$","2","2"),
  c("$\\beta_S$","2","2"),
  c("$\\text{Range}$"," -- ","1.5"),
  c("$\\text{Marginal variance}$"," -- ","0.5"),
  c("$\\xi_{com}$","-1","-1"),
  c("$\\sigma_{com}$","1","1"),
  c("$k_{com}$"," -- ","1"),
  c("$\\xi_{sci}$"," -- ","0"),
  c("$\\sigma_{sci}$"," -- ","0.8")
  ))

```

**Single-square simulations**

Two important variables may affect the accuracy of model outputs: the amount of commercial data and the number of fishing zones explored within one single declaration. In these simulations-estimations, we intend to explore the effect of these 2 variables.

First, the amount of commercial data; it is expected that when increasing the amount of data the accuracy of the predictions will be improved. In simulations, we progressively increase the number of fishing positions (10, 100 and 1000 fishing pings). As there are 10 pings per declaration, this means that the number of declarations increases respectively (1, 10 and 100 declarations).

Furthermore, the samples belonging to a single declaration can be sampled in a single restricted zone or can mix catches from distinct zones (within a unique statistical rectangle). Note that we make a distinction between fishing zones and statistical rectangles: fishing zones are included in a statistical rectangle and there can be several fishing zones within a statistical rectangles. Consequently, a declaration can mix catches that have been realized in zones of high biomass and in zones of low biomass. For instance, a declaration can aggregate data from several fishing operations that occured the same day within the same statistical rectangle, but that were realized on 2 distinct types of fishing grounds (and then on two distinct habitats). Reallocating uniformly the declaration will strongly homogeneise the actual catch, will mix up information from to different grounds and then may lead to a strong loss of accuracy in model outputs. It is expected that when the number of fishing zones within a declaration increases, the accuracy of the model outputs decreases. To assess the effect of such process, we simulated the pings of a fishing declaration assuming they were either realized in a single zone, in 3 distinct zones or in 5 distinct zones.

Concretely, for each fishing declaration the fishing pings are sampled in 2 steps: first the centroid of the zones are randomly sampled over the statistical rectangle (here the simulation domain) and then the fishing positions are randomly sampled within the radius of the centroid of the zones. The zone size was chosen to be a square of side 7. This is a realistic order of magnitude if we consider each zone is one fishing operation and that for each fishing operation the distance of the operation will not exceed 30km.

<!--  One fishing operation is about 3 or 4 hours at a speed of 4 nodes (7.4 km/h) so it makes a distance between 22 and 29 km. If we consider the side of the study domain (here a 25 x 25 grid) is $1° \approx 80km$ (as France is at 45° of latitude), a 7 cell zone is equivalent to a 22 km zone. -->

At each fishing position, a catch is realized conditionnally on the value of the latent field and follow the probability distribution described above ($\mathcal{L}_y$).

We compare three simulation/estimation configurations:

- a golden standard configuration: punctual observations ($Y_i$) are known exactly and the punctual-data model is fitted to the exact data.

- a configuration corresponding to the actual situation (uniform reallocation): simulated catches are summed into declarations and reallocated over the related fishing locations. We then fit the $Y_i^*$ model (the model fitted to punctual reallocated data).

- a configuration corresponding to our alternative approach: simulated catch are reallocated over fishing locations and we fit the $D_j$ model to the data (i.e. the one fitted at the level of the declarations data).

In addition to the metrics we introduced to compare model configuration ($MSPE$ and species-habitat parameter $\beta_S$), we also assess the simulated/estimated values for the intercept $\mu$, the observation variance parameter $\sigma^2$ and the zero-inflation parameter $\xi$.

**Multiple-square simulations**

In this case, we extend the study to more than one single rectangle, we simulate scientific data and include these in inference. Finally, in addition to the covariate, we also simulate/estimate a spatial random effect in the latent field as a Matérn GRF.

The study area is based on the case study; it includes the whole coast of the Bay of Biscay and cover several statistical rectangles (Figure \@ref(fig:MapSeveral)A). To tailor the case study, we simulate 3000 of fishing positions regrouped in 300 declarations (10 pings per declaration). Commercial data may not cover the full area, and consequently we allow the commercial samples to cover only 2/3 ot the area. Similarly to the single-square simulations, the sampling of the commercial points of a declaration are realized in 3 steps. (1) The declaration is randomly affected to one of the ICES rectangle. (2) The centroid of the fishing zone is randomly sampled within this statistical rectangle. (3) The 10 fishing punctual observations are randomly sampled within the fishing zone. The radius of the fishing zone is set so as the extent of a fishing operation does not exceed 30 km. Note that we do not explore the effect of exploring several zones within the same declaration as it is already done in the single-square simulations.

100 scientific samples are simulated following a random stratified plan; contrarly to commercial data they cover the entire study domain (Figure \@ref(fig:MapSeveral)A). Scientific observations are simulated following the observation equations of $\mathcal{L}_Y$ (with specific parameters for scientific data - Table 1).

We compare several model configurations:

- to assess what brings our alternative approach, we compare the $Y^*_i$ model and the $D_j$ model.

- to assess the information brought by each data source, we compare models built on scientific data only (scientific-based models), models built on commercial data only (commercial-based models) and models combining both data sources (integrated models).

In addition to the 2 metrics introduced at the beginning of the section ($MSPE$ and species-habitat parameter $\beta_S$), we also compare the simulated/estimated values of the range parameter.

Note that when including 2 datasets in the model, a difference in catchability should be accounted for and one of the two dataset will be considered as reference (see @alglave_combining_2022 for more details on parameterization). Here the reference level is set with scientific data and the parameter $k_{com}$ accounts for the relative ratio between scientific and commercial observation values. In simulations, $k_{com}$ is set to 1 so that both datasources are assumed to have the same catchability.

## Case-study: sole of the Bay of Biscay

To illustrate our method on a real case study, we applied the approach to the common sole of the Bay of Biscay. VMS-logbooks data were extracted for the bottom trawlers fleet (OTB). The methods to cross VMS-logbooks data and to filter the fleet is already extensively described in the previous papers [@alglave_combining_2022] and is not developped further here. Scientific data were extracted from the DATRAS database for the Orhago beam trawl survey [@biais_gerard_orhago_2003;@ices_report_2018]. To align the commercial and the scientific data, we filtered scientific data based on the minimum size of sole (24 cm for sole - @ices_report_2018-1). To illustrate the method, we compare the outputs of (1) the scientific-based model, (2) the integrated model fitted to scientific data and reallocated commercial data ($Y^*_i$ model) and (3) the integrated model fitted to scientific and declarations data ($D_j$ model).

The $D_j$ model faced convergence issues (some of the parameters were hardly estimated e.g. the range parameter). To ease convergence, we integrated in the analysis onboard observer data for the same fleet. They are considered as exact commercial data of catches. Integrating these data allow to have direct information on $Y_i$ and to better estimate the observation equations parameters (i.e. observation variance and zero-inflation parameter of commercial data).

Furthermore, we adopt a step estimation procedure to initialize the optimization algorithm for the $D_j$ model. We first fit the $Y^*_i$ model and use the estimates of the model to initialize the optimization of the $D_j$ model. We eventually fix the parameters that are hard to estimate in first steps of optimization (intercept $\mu$, covariate effect $\beta_S$, range and marginal variance) and let them free iteratively.

# Results

## Single-square simulations

```{r,warning=F}

load("res/Results_full_single_square.RData")

## Table of convergence
Results_conv <- Results
Results_conv$one <- 1
Results_conv_2 <- summaryBy(Convergence+one~
                              sequencesdepeche+
                              # zonespersequence+
                              aggreg_obs+
                              # b_true+
                              n_samp_com+
                              reallocation,
                            data=Results_conv,
                            FUN=sum) %>%
  dplyr::select(n_samp_com,
                sequencesdepeche,
                # zonespersequence,
                reallocation,
                aggreg_obs,
                Convergence.sum,
                one.sum) %>%
    mutate(aggreg_obs = ifelse(aggreg_obs == T,"Dj","Yi"),
           reallocation = ifelse(reallocation == 1,"Yes","No"),
           perc_convergence = round((1 - Convergence.sum / one.sum)*100,digits = 3)) %>%
  mutate(aggreg_obs = ifelse((aggreg_obs == "Yi" & reallocation == "Yes"),"Yi*",aggreg_obs))

Results_conv_2$n_samp_com <- as.integer(Results_conv_2$n_samp_com)
Results_conv_2$sequencesdepeche <- as.integer(Results_conv_2$sequencesdepeche)

Results_conv_2 <- Results_conv_2 %>%
  dplyr::rename(`Fishing positions` = n_samp_com,
                `Declarations` = sequencesdepeche,
                `Reallocation` = reallocation,
                `Likelihood level` = aggreg_obs,
                `Convergence (%)` = perc_convergence) %>%
  dplyr::select(-Convergence.sum,-one.sum)

```

```{r,warning=F}

## Plot Performance metric
Results <- Results %>%
  filter(b_true == 0 & Convergence == 0)
Results[,"RelBias_N"]=(Results[,"N_est"]-Results[,"N_true"])/Results[,"N_true"]
Results[,"RelBias_beta"]=(Results[,"beta1_est"]-Results[,"beta1_true"])/Results[,"beta1_true"]
Results[,"Bias_b"]=(Results[,"b_est"]-Results[,"b_true"]) / ifelse(Results[,"b_true"] != 0,Results[,"b_true"],1)
Results[,"Bias_sigma_com"]=(Results[,"sigma_com_est"]-Results[,"sigma_com_true"]) / Results[,"sigma_com_true"]
Results[,"Bias_q1_com"]=(Results[,"q1_com_est"]-Results[,"q1_com_true"]) / Results[,"q1_com_true"]
Results[,"Bias_intercept"]=(Results[,"intercept_est"]-Results[,"intercept_true"]) / Results[,"intercept_true"]

Results <- Results %>%
  mutate(relloc_aggreg = paste0("Reallocation: ",ifelse(reallocation==0,"No","Yes")," -  Likelihood: ",ifelse(aggreg_obs==T,"Dj","Yi"))) %>%
  mutate(relloc_aggreg = ifelse(aggreg_obs == F & reallocation == 1,"Reallocation: Yes -  Likelihood: Yi*",relloc_aggreg))

Results$zonespersequence <- as.factor(Results$zonespersequence)
Results$b_true <- as.factor(Results$b_true)
Results$relloc_aggreg <- factor(Results$relloc_aggreg,
                                levels = c("Reallocation: No -  Likelihood: Yi",
                                           "Reallocation: Yes -  Likelihood: Yi*",
                                           "Reallocation: Yes -  Likelihood: Dj"))

Results_2 <- Results %>%
  filter(n_samp_com > 10)

MSPE_S_plot <- ggplot()+
  geom_boxplot(data = Results_2,
               aes(x = zonespersequence,
                   y = mspe,
                   fill = relloc_aggreg),
               outlier.colour = NA)+
  theme_bw()+
  facet_wrap(.~factor(n_samp_com))+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1)+
  ylab("MSPE")+
  xlab("")+
  scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                labels = trans_format("log10", math_format(10^.x)))+
  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
                                "Reallocation: Yes -  Likelihood: Yi*",
                                "Reallocation: Yes -  Likelihood: Dj"),
                     values=c("gold","#F8766D","#00BA38"))

Beta_plot <- ggplot()+
  geom_boxplot(data = Results_2,
               aes(x = zonespersequence,
                   y = beta1_est,
                   fill = relloc_aggreg),
               outlier.colour = NA)+
  geom_hline(yintercept = 0,linetype="dashed")+
  geom_hline(yintercept = 2,col="red",linetype="dashed")+
  theme_bw()+
  facet_wrap(.~factor(n_samp_com))+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1)+
  # ylim(-1,1)+ 
  ylab(parse(text = TeX('$\\beta_S$')))+
  xlab("")+
  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
                                "Reallocation: Yes -  Likelihood: Yi*",
                                "Reallocation: Yes -  Likelihood: Dj"),
                     values=c("gold","#F8766D","#00BA38"))

RelBias_N_plot <- ggplot()+
  geom_boxplot(data = Results_2,
               aes(x = zonespersequence,
                   y = RelBias_N,
                   fill = relloc_aggreg),
               outlier.colour = NA)+
  geom_hline(yintercept = 0,linetype="dashed")+
  theme_bw()+
  facet_wrap(.~factor(n_samp_com))+
  theme(legend.title = element_blank(),
        aspect.ratio = 1,
        legend.position = 'none')+
  ylim(-1,1)+
  ylab("Relative bias of biomass")+
  xlab("")+
  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
                                "Reallocation: Yes -  Likelihood: Yi*",
                                "Reallocation: Yes -  Likelihood: Dj"),
                     values=c("gold","#F8766D","#00BA38"))

legend <- as_ggplot(cowplot::get_legend(MSPE_S_plot+theme(legend.position="right",legend.title = element_blank())))

Perf.metric_single_square <- plot_grid(MSPE_S_plot,
                                       Beta_plot,
                                       # RelBias_N_plot,
                                       legend,
                                       ncol = 1,
                                       rel_heights = c(1,1,0.25))

ggsave("images/Perf.metric_single_square.png",width = 5*1.25, height = 6.5*1.25)

Results_3 <- Results_2 %>%
  filter(n_samp_com == 1000)

q1_plot <- ggplot()+
  geom_boxplot(data = Results_3,
               aes(x = relloc_aggreg,
                   y = q1_com_est,
                   fill = relloc_aggreg),
               outlier.colour = NA)+
  geom_hline(yintercept = 0,linetype="dashed")+
  geom_hline(yintercept = -1,col="red",linetype="dashed")+
  theme_bw()+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1,
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
                               "Reallocation: Yes -  Likelihood: Yi*",
                               "Reallocation: Yes -  Likelihood: Dj"),
                    values=c("gold","#F8766D","#00BA38"))+
  xlab("")+ylab(parse(text = TeX('$\\xi$')))+
  ggtitle("Zero-inflation parameter")

sigma_plot <- ggplot()+
  geom_boxplot(data = Results_3,
               aes(x = relloc_aggreg,
                   y = sigma_com_est,
                   fill = relloc_aggreg),
               outlier.colour = NA)+
  geom_hline(yintercept = 0,linetype="dashed")+
  theme_bw()+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1,
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
                               "Reallocation: Yes -  Likelihood: Yi*",
                               "Reallocation: Yes -  Likelihood: Dj"),
                    values=c("gold","#F8766D","#00BA38"))+
  xlab("")+ylab(parse(text = TeX('$\\sigma^2$')))+
  ggtitle("Observation variance parameter")+
  geom_hline(yintercept = 1,col="red",linetype="dashed")

intercept_plot <- ggplot()+
  geom_boxplot(data = Results_3,
               aes(x = relloc_aggreg,
                   y = intercept_est,
                   fill = relloc_aggreg),
               outlier.colour = NA)+
  geom_hline(yintercept = 0,linetype="dashed")+
  theme_bw()+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1,
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(breaks = c("Reallocation: No -  Likelihood: Yi",
                               "Reallocation: Yes -  Likelihood: Yi*",
                               "Reallocation: Yes -  Likelihood: Dj"),
                    values=c("gold","#F8766D","#00BA38"))+
  xlab("")+ylab(parse(text = TeX('$\\mu$')))+
  ggtitle("Intercept of the latent field")+
  geom_hline(yintercept = 2,col="red",linetype="dashed")

par_plot <- plot_grid(q1_plot,sigma_plot,intercept_plot,nrow = 1,align = "hv")

legend <- as_ggplot(cowplot::get_legend(q1_plot+theme(legend.position="bottom",legend.title = element_blank())))

par_plot <- plot_grid(par_plot,legend,rel_heights = c(1,0.08),ncol=1)

ggsave("images/par_plot_single_square.png",width = 10.5, height = 4)

```

Reallocation has a major effect on predictions and estimates accuracy (Figure \@ref(fig:PerfMetricSingle)). Reallocating data conduct to a 10 to 200 times decrease in accuracy for spatial predictions when fitting the model to reallocated catch (MSPE gold compared to red boxplots). Accuracy decreases as the number of visited zones within a declaration increases. Besides, $\beta_S$ estimates are biased and reallocation leads to the loss of the species-habitat relationship as the number of fishing zones (within a declaration) increases ($\beta_S$ estimates get closer to 0). Increasing the number of samples does not improve inference. Regarding other parameters estimates, the zero-inflation parameter ($\xi$) is over-estimated (i.e. when uniformly reallocating commercial data, the quantity of positive observations is under-estimated), the observation variance ($\sigma$) is underestimated (i.e. the data is estimated to be less noisy than they actually are) and the intercept of the latent field ($\mu$) is slightly over-estimated (Figure \@ref(fig:ParBiasSingle)).

The $D_j$ model allows to recover the species-habitat relationship and to improve the accuracy of the spatial predictions (Figure \@ref(fig:PerfMetricSingle)). Still, the $D_j$ model outputs are not as accurate as the ones of the golden-standard. Furthermore, the zero-inflation parameter is unbiased when the model is fitted to catch declarations $D_j$. Other parameters (observation variance, intercept) are also better estimated even though they remain slightly biased (Figure \@ref(fig:ParBiasSingle)). This alternative model have some convergence difficulties (Table 2) as 8% of the model did not converged when sample size is medium (100 pings) and only 3% did not when sample size is large (1000 pings).

![(\#fig:PerfMetricSingle) Performance metric for single-square simulations. Columns: number of commercial fishing positions. x-axis: number of zones visited within each declaration. 'Reallocation: ', data are or are not reallocated in simulations. 'Likelihood: ', the likelihood is computed on exact punctual observations $Y_i$, on reallocated observations $Y_i^*$ or on catch declarations $D_j$. Gold: golden standard. Red: uniform reallocation ($Y_i^*$ model). Green: model-based reallocation ($D_j$ model). Simulations conducted on 10 fishing positions are not represented as they mostly did not converged for the $D_j$ model (see Table 2).](images/Perf.metric_single_square.png){width=75%}

![(\#fig:ParBiasSingle) Parameters relative bias for single-square simulations. 'Reallocation: ', data are or are not reallocated in simulations. 'Likelihood: ', the likelihood is computed on exact punctual observations $Y_i$, on reallocated observations $Y_i^*$ or on catch declarations $D_j$. Gold: golden standard. Red: uniform reallocation ($Y_i^*$ model). Green: model-based reallocation ($D_j$ model). Only the simulations with 1000 fishing positionss are represented. Black line: zero value. Red line: parameter true value.](images/par_plot_single_square.png)

```{r}

knitr::kable(Results_conv_2,booktabs = T,align = "c",
                  caption = "Single-square simulations - Percentage of convergence per simulation-estimation configuration.")

```

\newpage

## Multiple-square simulations

```{r,warning=F}

load("res/Results_full_multi_square.RData")

Results_conv <- Results_2
Results_conv$one <- 1

Results_conv_2 <- summaryBy(converge+one~
                              Model+
                              aggreg_obs,
                            data=Results_conv,
                            FUN=sum) %>%
  dplyr::select(Model,
                aggreg_obs,
                converge.sum,
                one.sum) %>%
    mutate(aggreg_obs = ifelse(aggreg_obs == T,"Dj","Yi"),
           perc_convergence = round((1 - converge.sum / one.sum)*100,digits = 3))

Results_conv_2$aggreg_obs[which(Results_conv_2$aggreg_obs == "Yi")] <- "Yi*"

Results_conv_2 <- Results_conv_2 %>%
  dplyr::rename(`Likelihood level` = aggreg_obs,
                `Convergence (%)` = perc_convergence) %>%
  dplyr::select(-converge.sum,-one.sum)

Results_conv_2$Model <- factor(Results_conv_2$Model,levels = c("Scientific model", "Commercial model", "Integrated model"))

Results_conv_2 <- Results_conv_2[-which(Results_conv_2$Model == "Scientific model" &
                                          Results_conv_2$`Likelihood level` == "Dj"),]
Results_conv_2$`Likelihood level`[which(Results_conv_2$Model == "Scientific model" &
                       Results_conv_2$`Likelihood level` == "Yi*")] <- " "

```

```{r,warning=F}

Results_plot <- Results_2 %>%
  filter(converge == 0) %>%
  filter(simu_type == "Unsampled Rectangles")

Results_plot[,"RelBias_N"]=(Results_plot[,"N_est"]-Results_plot[,"N_true"])/Results_plot[,"N_true"]
Results_plot[,"RelBias_N.2"]=(Results_plot[,"N_est.2"]-Results_plot[,"N_true.2"])/Results_plot[,"N_true.2"]
Results_plot[,"RelBias_beta"]=(Results_plot[,"beta1_est"]-Results_plot[,"beta1_true"])/Results_plot[,"beta1_true"]

Results_plot$obs <- NA
Results_plot$obs[which(Results_plot$aggreg_obs == T)] <- "Dj"
Results_plot$obs[which(Results_plot$aggreg_obs == F)] <- "Yi*"
Results_plot$obs[which(Results_plot$Estimation_model == 2)] <- "Scientific data"
Results_plot$obs <- factor(Results_plot$obs,levels = c("Yi*","Dj","Scientific data"))

Results_plot$Model[which(Results_plot$Model == "Scientific model")] <- "Scientific"
Results_plot$Model[which(Results_plot$Model == "Commercial model")] <- "Commercial"
Results_plot$Model[which(Results_plot$Model == "Integrated model")] <- "Integrated"

# Results_plot_fullArea <- Results_plot
# Results_plot_UnsampArea <- Results_plot

beta_plot <- ggplot()+
  geom_boxplot(data = Results_plot,
               aes(x = Model,
                   y = beta1_est,
                   fill = obs))+
  geom_hline(yintercept = 0,linetype="dashed")+
  theme_bw()+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1)+
  # facet_wrap(.~simu_type)+
  ylab(TeX("$\\beta_S$")) +
  xlab("")+
  geom_hline(yintercept = 2,col="red",linetype="dashed")+
  scale_color_manual(breaks = c("Scientific model",
                                "Yi*",
                                "Dj"),
                     values=c("#619CFF","#F8766D","#00BA38"))

MSPE_S_plot <- ggplot()+
  geom_boxplot(data = Results_plot,
               aes(x = Model,
                   y = mspe,
                   fill = obs))+
  theme_bw()+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1)+
  # facet_wrap(.~simu_type)+
  ylab("MSPE")+xlab("")+
  scale_color_manual(breaks = c("Scientific model",
                                "Yi*",
                                "Dj"),
                     values=c("#619CFF","#F8766D","#00BA38"))+
  ylim(0,NA)

Range_plot <- ggplot()+
  geom_boxplot(data = Results_plot,
               aes(x = Model,
                   y = Range,
                   fill = obs))+
  geom_hline(yintercept = 0,linetype="dashed")+
  theme_bw()+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1)+
  # facet_wrap(.~simu_type)+
  ylab("Range")+
  xlab("")+
  geom_hline(yintercept = 0.6,col="red",linetype="dashed")+
  scale_color_manual(breaks = c("Scientific model",
                                "Yi*",
                                "Dj"),
                     values=c("#619CFF","#F8766D","#00BA38"))+
  ylim(-0.1,2)

legend <- as_ggplot(cowplot::get_legend(MSPE_S_plot+theme(legend.position="bottom",legend.title = element_blank())))

pl_plot <- plot_grid(MSPE_S_plot,
                     beta_plot,
                     Range_plot,
                     nrow = 1,
                     align="hv")

Perf.metric_multi_square <- plot_grid(pl_plot,
                                      legend,
                                      ncol = 1,
                                      rel_heights = c(1,0.15))

ggsave("images/Perf.metric_multiple_square.png",width = 12, height = 4)

# load("res/lkl_prof_beta.RData")

```

Whether the model is fitted at the punctual level or at the declaration level, the contribution of either scientific or commercial data can be clearly evidenced from the MSPE plot: the errors related to the integrated model are smaller than the single-data models. This can be well illustrated from Figure \@ref(fig:MapSeveral). Integrating scientific and commercial data allows to (1) capture the hotspot missed by commercial data through scientific data and (2) better capture the local correlation structures through the dense commercial data.

Furthermore, consistently with single-square simulations, uniform reallocation conducts to a loss in both the predictions accuracy and the species-habitat relationship (Figure \@ref(fig:PerfMetricSeveral)) compared to the model built on commercial declarations ($D_j$).

Interestingly, in addition to the species-habitat relationship, uniform reallocation also affects the spatial autocorrelation terms such as the range parameter. The model fitted on reallocated data ($Y_i^*$) provides biased range estimates while the $D_j$ model provides unbiased estimates. This is a consequence of the loss of the species-habitat relationship: when uniformly reallocating declarations, part of the variability related to the covariate effect is captured by the random effect. Consequently, the range parameter captures both the autocorrelation related to the actual random effect and to the covariate. The $D_j$ model allows to recover and disentangle the effect of the species-habitat relationship and of the random effect. This is evidenced in Figure \@ref(fig:MapSeveral) where the model fitted to reallocated catch $Y_i^*$ provides smoothed maps and does not capture the relatively small scale patterns that are shaped by the covariate. On the other hand, the $D_j$ model (as the scientific-based model) better captures and disentangles the covariate effect and the spatial random effect and then provides predictions that better fit to the small-scale patterns of the species distribution. However, this goes with some difficulty in convergence as only 75% of the model built on catch declarations converge (Table 3).

![(\#fig:PerfMetricSeveral) Performance metric for the multiple-square simulations. Red line: true value for the range and the species-habitat parameter ($\beta_S$). Red: uniform reallocation ($Y_i^*$ model). Green: model-based reallocation ($D_j$ model). Blue: scientific-based model.](images/Perf.metric_multiple_square.png)

![(\#fig:MapSeveral) Relative distribution of simulated/estimated biomass field. A: Simulated biomass field with scientific samples (red) and statistical rectangles. The rectangles that have not been sampled by commercial data are the transparent rectangles. They represent 1/3 of the full area. B: simulated biomass field. C: biomass field from the scientific-based model. $Y_i^*$: $Y_i^*$ model (D, E). $D_j$: $D_j$ model (F, G). Scientific model: model fitted to scientific data only. Commercial model: model fitted to commercial data only. Integratedd model: model fitted to both data sources.](images/Map_multi_square.png)

```{r}

knitr::kable(Results_conv_2,booktabs = T,align = "c",
                  caption = "Multiple-square simulations - Percentage of convergence per simulation-estimation configuration.")

```

<!-- see the file 'Scripts/In_progress/plot_multiple_square_2.R' -->

## Real case study

```{r}

## Load data
load("res/sci_df.RData")

load("res/realloc_rest_df.RData")

load("res/no.realloc_df.RData")

load("res/est_par_df_full_3.RData")

## Plot maps
no.realloc_plot <- ggplot(no.realloc_df)+
  geom_point(aes(x=x,y=y,col=S_x),shape=15,size=2)+
  scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
  ggtitle("Integrated model",subtitle = "Commercial likelihood on Yi*")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))+
  geom_sf(data = mapBase)+
  coord_sf(xlim = c(-6,0), ylim = c(43,48+0.25), expand = FALSE)+
  xlab("")+ylab("")

realloc_rest_plot <- ggplot(realloc_rest_df)+
  geom_point(aes(x=x,y=y,col=S_x),shape=15,size=2)+
  scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
  ggtitle("Integrated model",subtitle = "Commercial likelihood on Dj")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))+
  geom_sf(data = mapBase)+
  coord_sf(xlim = c(-6,0), ylim = c(43,48+0.25), expand = FALSE)+
  xlab("")+ylab("")

sci_plot <- ggplot(sci_df)+
  geom_point(aes(x=x,y=y,col=S_x),shape=15,size=2)+
  scale_color_distiller(palette = "Spectral",limits=c(0,NA))+
  ggtitle("Scientific model")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))+
  geom_sf(data = mapBase)+
  coord_sf(xlim = c(-6,0), ylim = c(43,48+0.25), expand = FALSE)+
  xlab("")+ylab("")

# plot(log(no.realloc_df$S_x),log(sci_df$S_x))
# cor(log(no.realloc_df$S_x),log(sci_df$S_x),method = "spearman")
# plot(log(realloc_rest_df$S_x),log(sci_df$S_x))
# cor(log(realloc_rest_df$S_x),log(sci_df$S_x),method = "spearman")

case_study_plot <- plot_grid(sci_plot,no.realloc_plot,realloc_rest_plot,ncol=3,align="hv")

ggsave("images/case_study_plot.png",width = 12, height = 4)

## Parameter estimates
est_par_df_full_3$lkl[which(est_par_df_full_3$lkl == "Yi")] <- "Yi*"
est_par_df_full_3$lkl <- factor(est_par_df_full_3$lkl,levels = c("Dj","Yi*","Scientific model"))
est_par_df_full_3 <- est_par_df_full_3 %>%
  filter(par_names != "substr_Sand_Coarse_substrate")
par_plot <- ggplot(est_par_df_full_3, aes(y=par_val, x=par_names))+
  geom_point(
    aes(color = lkl),
    position = position_dodge(0.5),
    size=2
  )+
  geom_errorbar(aes(ymin = CI.inf, ymax = CI.sup, color = lkl),
                position = position_dodge(0.5),width=0.4
  )+
  scale_color_manual(breaks = c("Scientific model",
                                "Yi*",
                                "Dj"),
                     values=c("#619CFF","#F8766D","#00BA38"))+
  geom_hline(yintercept=0, linetype="dashed", color = "red",alpha=0.4)+
  xlab("")+ylab("")+
  theme_bw()+
  theme(legend.title = element_blank(),
        legend.position = "none",
        aspect.ratio = 1)+
  coord_flip()+
  scale_x_discrete(labels=c('k_com'=parse(text = TeX('$k_{com}$')),
                            'Sigma_com'=parse(text = TeX('$\\sigma_{com}$')),
                            'q1_com'=parse(text = TeX('$\\xi_{com}$')),
                            'Sigma_sci'=parse(text = TeX('$\\sigma_{sci}$')),
                            'q1_sci'=parse(text = TeX('$\\xi_{sci}$')),
                            'Range'="Range",
                            'MargSD'="Marginale variance",
                            'substr_Mud_sediment'=parse(text = TeX('$\\beta_{S}$')),
                            'intercept'=parse(text = TeX('$\\mu$'))))

legend <- as_ggplot(cowplot::get_legend(
  par_plot+
    theme(legend.position="right",
          legend.title = element_blank()
          )))

par_plot <- plot_grid(par_plot,legend,ncol=1,rel_heights = c(1,0.25),align="hv")

ggsave("images/par_plot.png",width = 7.5, height = 6)

```

Consistently with simulations, the $D_j$ model modifies some of the parameters estimates and revises the spatial pattern of the species distribution compared with the $Y^*_i$ model. In particular, the substrate effect is recovered in the $D_j$ model and fall in the same range as the scientific-based estimate (Figure \@ref(fig:CaseStudyMap)). The zero-inflation parameter $\xi$ is revised downwards (i.e. there are actually more zero-values than in the reallocated data) while the observation variance of commercial data is revised upwards (i.e. the commercial data are noisier than expected when fitting the model on $Y_i^*$).

In addition, uncertainty is also revised when fitting the model at the declaration level. For instance, when comparing the $Y^*_i$ model to the $D_j$ model, the confidence intervals of $\beta_S$, the marginal variance, the range, $\xi_{com}$, $\sigma_{com}$ are much wider. This emphasizes that uncertainty is probably underestimated in the $Y^*_i$ model compared with the $D_j$ model. Now, when comparing the scientific-based model and the integrated $D_j$ model, some parameters are better estimated than when only scientific data feed the model. For instance, while in the scientific-based model the substrate effect was not significant, in the integrated model built on $D_j$ substrate is significant and the confidence interval is smaller.

On the contrary, other parameters do not seems be well estimated in either the $Y_i^*$ or the $D_j$ models. For instance, compared to the scientific-based model, the intercept is revised upwards when building the likelihood on $Y_i$ and revised downards when working on $D_j$. This is consistent with simulations results, see Figure \@ref(fig:ParBiasSingle).

Regarding the maps of the species distribution, fitting the model at the $D_j$ level strongly modifies the model biomass field compared with the $Y_i^*$ model. In particular, the substrate covariate have a sharper effect on species distribution and the intensity of the hotspots are revised when fitting the model on $D_j$.

Finally, the $D_j$ model fitted only on commercial data does not converge (while the one fitted on $Y_i^*$ does) emphasizing the model fitted on catch declarations face difficulties to converge and require punctual observations (here survey data and on-board observer data) to converge on real data.

![(\#fig:CaseStudyPar) Parameters obtained with scientific-based model, the integrated model fitted on reallocated catch $Y_i^*$ and the integrated model fitted on catch declarations $D_j$.](images/par_plot.png)

![(\#fig:CaseStudyMap) Maps obtained from (left) the scientific-based model, (center) the integrated model fitted on punctual reallocated catch $Y_i^*$, (right) the integrated model fitted on catch declarations $D_j$.](images/case_study_plot.png)

# Discussion

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

# Supplementary material

**Notations**

We model catch declarations $D_j$ (available at coarse resolution through logbooks data) as a sum of $Y_{i}$ punctual observations (which are unknown i.e. latent variables) each one realised at one fishing position $x_{i}$ (known through VMS data).

We note:

- $\mathcal{P}_j$ : the vector of all fishing positions related to the $j^{th}$ declaration.

- $j \in \{1,...,l\}$ with $l$ the number of declarations.

- $i \in \{1,...,m_j\}$ with $m_j$ the number of fishing positions belonging to the $j^{th}$ declaration.

$$D_j=\sum_{i \in \mathcal{P}_j}{Y_{i}}$$

**Reparameterization of the lognormal distribution**

The lognormal distribution is usually written as $Z \sim \operatorname{L}(\rho;\sigma^2)$ with $Z=e^{\rho+\sigma N}$ and $N \sim \mathcal{N}(0,1)$. In this case, $E(Z)=e^{\rho + \frac{\sigma^2}{2}}$ and $Var(Z)=(e^{\sigma^2}-1)e^{2 \rho + \sigma^2}$.

We choose to slightly reparameterize the lognormal distribution. Let's define $\rho = ln(\mu) - \frac{sigma^2}{2}$, then:

- $Z=\mu e^{\sigma N - \frac{sigma^2}{2}}$

- $E(Z)=\mu$

- $Var(Z)=\mu^2(e^{\sigma^2} - 1) \Leftrightarrow	\sigma^2=ln(\frac{Var(Z)}{E(Z)^2} + 1)$

**$D_j$ probability distribution and moments**

We have to express the probability distribution of $D_j$ and its moments as a function of $Y_{i}$ and its related moments. Let's assume $Y_{i} = C_{i}.Z_{i}$ is a zero-inflated lognormal distribution with $C_{i}$ a binary random variable and $Z_{i}$ a lognormal random variable.

$$C_{i} \sim \mathcal{B}(1-p_{i})$$ 
with $p_{i}=exp(-e^\xi .S(x_{i}))$ the probability to obtain a zero value.

$$Z_{i} \sim \operatorname{L}(\frac{S(x_{i})}{1-p_{i}},\sigma^2)$$

Here, $Y_{i}$, $C_{i}$ and $Z_{i}$ are observations of a latent field $S(x_{i})$ at a sampled point $x_{i}$.

**Probability of obtaining a zero declaration**

\begin{align*}
P(D_j = 0 \vert S, X) & = \prod_{i\in \mathcal{P}_j} P(Y_{i} = 0 \vert S, X),\nonumber \\
                      & = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j.
\end{align*}

**Expectancy of a positive declaration**

Following calculations are supposed to be conditionnal on $S$ and $X$.

$$E(D_j\vert D_j >0) =  \sum_{i\in \mathcal{P}_j} E(C_{i} Z_{i}\vert \exists i\in\mathcal{P}_j , C_{i}=1)$$

\begin{align*}
E(D_j\vert D_j > 0) & = E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } )  / P\left ( D_j > 0\right ),   \\
& = E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } )  / \left (1-\pi_j\right).   \\
\end{align*}

As $E(D_j 1_{ \left \lbrace D_j > 0\right\rbrace } ) = E(D_j ),$

\begin{align*}
E(D_j\vert D_j > 0) & = \left (1- \pi_j\right)^{-1} E(D_j)   ,   \nonumber \\
& = \left (1- \pi_j\right)^{-1} \sum_{i\in \mathcal{P}_j} E(C_{i} Z_{i}),\nonumber \\
& = \left (1- \pi_j\right)^{-1} \sum_{i\in \mathcal{P}_j} (1-p_{i}) \frac{S(x_{i})}{1-p_{i}}, \nonumber \\
& = \left (1- \pi_j\right)^{-1}\sum_{i\in \mathcal{P}_j} S(x_{i}).
\end{align*}

**Variance of a positive declaration**

$$Var(D_j \vert D_j >0)  = E(D_j^2 \vert D_j >0)- E(D_j \vert D_j >0)^2.$$

$$E(D_j^2 \vert D_j >0) = (1-\pi_j)^{-1} E(D_j^2 1_{\left \lbrace D_j >0\right\rbrace}) = (1-\pi_j)^{-1} E(D_j^2 )$$

$$E(D_j \vert D_j >0)^2 = ((1-\pi_j)^{-1} E(D_j 1_{\left \lbrace D_j >0\right\rbrace}))^2 =  (1-\pi_j)^{-2} E(D_j)^2$$

Then,

$$Var(D_j \vert D_j >0) = (1-\pi_j)^{-1} E(D_j^2 ) - (1-\pi_j)^{-2} E(D_j)^2 = (1-\pi_j)^{-1} Var(D_j) - \frac{\pi_j}{(1-\pi_j)^2} E(D_j)^2.$$

As the $(Y_{i})_{i\in \mathcal{P}_j}$ are independent, $Var(D_j) =\sum_{i \in\mathcal{P}_j} Var(Y_{i}) = \sum_{i \in\mathcal{P}_j} Var(C_{i}.Z_{i})$.

\begin{align*}
Var(C_{i} Z_{i}) & = E( C_{i}^2 Z_{i}^2 ) - E(C_{i} Z_{i})^2,\\
& =  E( C_{i}^2) E(Z_{i}^2 ) -E(C_{i})^2 E(Z_{i})^2,\\
& = (1-p_{i}) E(Z_{i}^2 ) - (1-p_{i})^2 E(Z_{i})^2,\\
& = (1-p_{i})(Var(Z_{i})+E(Z_{i})^2)-(1-p_{i})^2E(Z_{i})^2,\\
& = \frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-1)+\frac{S(x_{i})^2}{1-p_{i}}-S(x_{i})^2,\\
& = \frac{S(x_{i})^2}{1-p_{i}}(e^{\sigma^2}-(1-p_{i}))
\end{align*}

**Sum up of the main formulas**

$$P(D_j = 0 \vert S, X) = \exp{ \left \lbrace- \sum_{i\in \mathcal{P}_j} e^{\xi}. S(x_{i})\right \rbrace} = \pi_j$$

$$E(D_j \vert D_j > 0)=\frac{\sum_{i \in \mathcal{P}_j} S(x_{i})}{1-\pi_j}$$

$$Var(D_j \vert D_j > 0) = \frac{\sum_{i \in \mathcal{P}_j} Var(Y_{i})}{1-\pi_j} - \frac{\pi_j}{(1-\pi_j)^2}E(D_j)^2$$

$$Var(Y_{i})=\frac{S(x_{i})^2}{1-p_{j}}(e^{\sigma^2}-(1-p_{i}))$$

Assuming $D_j \vert D_j > 0$ also follows a lognormal distribution we can write:

$$D_j \vert D_j > 0 \sim \operatorname{L}( \mu_j = E(D_j \vert D_j > 0), \sigma_j^2= ln(\frac{Var(D_j \vert D_j > 0)}{E(D_j \vert D_j > 0)^2} + 1))$$